Grappa: Scalable Computing on Mass-Market Hardware for Irregular, Low-Locality Applications

### 2012-07-22 doc/asplos13/evaluation.tex
include uts plots, xmt uts temp plot: todo overlay

\subsubsection{Unbalanced tree search}

We ran UTS-mem on Grappa and the XMT with a binomial and two geometric
100M-vertex trees and a geometric 1.6B-vertex tree
\footnote{T3L: -t 0 -b 2000 -q 0.200014 -m 5 -r 7\\
          T2L: -t 1 -a 2 -d 23 -b 7 -r 220\\
          T1L: -t 1 -a 3 -d 13 -b 4 -r 29\\
          T1XL: -t 1 -a 3 -d 15 -b 4 -r 29\\}.
Both systems do relatively poorly on the binomial tree because it is very deep and a
traversal produces little parallelism, which both systems depend on.
\TODO{data from our trace would be nice here}. Figure~\ref{fig:grappa-xmt-uts}
shows the performance in terms of number of vertices visited per
second versus number of compute nodes. The Grappa results are for the
best parameter values from a limited search over \flushtimeout~and
\asyncforthr~\TODO{want to say this in general at beginning of eval?}.
Grappa with 16 machines is faster than the entire XMT. Grappa achieves
\checkme{112Mvert/s} with \checkme{60 machines} and the
XMT achieves only 50Mvert/s. Scaling up, Grappa adds 2Mvert/s/nodes
and the XMT adds 850Kvert/s/node. Grappa keeps scaling up through 128
machines, although scaling declines because of the unscalability of our aggregation
mechanism as number of network endpoints increases.

The XMT plataeus around \checkme{60} processors for 100M, 1.6B, and
4.2B-vertex trees, so performance is not limited by availability of
parallelism for this implementation of UTS-mem.

{\em you can rewrite this or blow it off -- just trying to get some thoughts in before leaving}
Despite our efforts to tune the UTS implementation specific to Cray XMT, performance does not scale well with increasing processor count, flattening out around sixty processors.  When we increase the size of the tree, we find that performance does not improve, suggesting that performance is not limited by task parallelism.   Cray's performance tools show that an increasing number of memory retry operations are generated by code within the runtime system.  Retries are performed by the memory controller when remote synchronization operations fail to find the full-bit associated with each memory location in the unavailable state.  Retries are issued at low priority relative to new memory operations issued by the processor by other contexts, so they consume what would otherwise be unused injection bandwidth.  On a full-bandwidth system such as the MTA-2, retries have no impact on the progress of tasks other than their own.  On a Cray XMT, network bandwidth is limited, so retries create congestion.  In comparison, Grappa performs synchronization without retries, delaying response at the receiving end until ready to notify the sender to proceed.  This saves bandwidth and permits scaling of tasks performing synchronization even on low injection rate networks.

\TODO{why is XMT getting different results for T1L,T2L; would expect it to be the machine with the most uniform results.}

\TODO{sensitivity study for \asyncforthr parameter, ie caching amount. Although we can take advantage of locality in children array and the XMT cannot, we still win without it}

\begin{figure}[h]
    \begin{center}
      \includegraphics[width=0.5\textwidth]{figs/grappa-xmt-uts.pdf}
    \end{center}
    \caption{describe...}
    \label{fig:grappa-xmt-uts}
\end{figure}

@@

@@

>>>>

\subsubsection{Unbalanced tree search in memory}
We ran UTS-mem on Grappa and the XMT with a binomial and two geometric
100M-vertex trees (T3L,T1L,T2L), a geometric 1.6B-vertex tree (T1XL), and a geometric
4.2B-vertex tree (T1XXL).
Both systems do relatively poorly on the binomial tree because it is very deep and a
traversal produces too little parallelism to get good bandwidth.
Figure~\ref{fig:grappa-xmt-uts}
shows the performance in terms of number of vertices visited per
second versus number of compute nodes. The Grappa results shown are for 5
cores per node and the best parameter values from a search over
runtime and application parameters.
Grappa with \checkme{16} machines is faster than the entire XMT. Grappa achieves
\checkme{188Mvert/s} with \checkme{128 machines} and the
XMT achieves only 50Mvert/s, plateauing at 60 nodes. Beyond 90 nodes,
Grappa adds 1.4 Mvert/s/node. The XMT scales at 850 Kvert/s/node,
until it plateaus. Grappa keeps scaling up through 128---the maximum we
could use---although scaling declines because of the unscalability of our aggregation
mechanism as number of network endpoints increases. Using 6 cores per
node gave Grappa up to a 20\% benefit at each data point, but Grappa could
not run well beyond 90 machines with this core configuration.

%% UTS: performance comparison
\begin{figure}[ht]
    \begin{center}
      \includegraphics[width=0.5\textwidth]{figs/uts_compare.pdf}
    \end{center}
    \caption{Performance of in-memory unbalanced tree search. Grappa
        keeps scaling through 128 nodes and beats the XMT with 16
        nodes.}
    \label{fig:uts_compare}
\end{figure}
\begin{figure}[ht]
    \begin{center}
      \includegraphics[width=0.5\textwidth]{figs/uts_xmt_temp.pdf}
    \end{center}
    \caption{This is XMT uts result, \TODO{overlay the plots}}
    \label{fig:uts_compare_temp}
\end{figure}

Despite our efforts to tune the UTS implementation specific to Cray
XMT, performance does not scale well with increasing processor count,
flattening out around sixty processors.  When we increase the size of
the tree from 100M to 4.2B, we find that performance does not improve, suggesting that performance is not limited by task parallelism.   Cray's performance tools show that an increasing number of memory retry operations are generated by code within the runtime system.  Retries are performed by the memory controller when remote synchronization operations fail to find the full-bit associated with each memory location in the unavailable state.  Retries are issued at low priority relative to new memory operations issued by the processor by other contexts, so they consume what would otherwise be unused injection bandwidth.  On a full-bandwidth system such as the MTA-2, retries have no impact on the progress of tasks other than their own.  On a Cray XMT, network bandwidth is limited, so retries create congestion.  In comparison, Grappa performs synchronization without retries, delaying response at the receiving end until ready to notify the sender to proceed.  This saves bandwidth and permits scaling of tasks performing synchronization even on low injection rate networks.

\TODO{ we still win with th=1 }

@@

%% UTS: chunk size
\begin{figure}[ht]
    \begin{center}
      \includegraphics[width=0.5\textwidth]{figs/uts_chunksize.pdf}
    \end{center}
    \caption{Performance of UTS-Mem with varying maximum chunk size of
    steals, run with 30 nodes, 6 cores per node, 4000 workers,
    \checkme{6M flush ticks}}
    \label{fig:uts_chunksize}
\end{figure}

@@

% uts threshold
\begin{figure}[ht]
    \begin{center}
      \includegraphics[width=0.5\textwidth]{figs/uts_threshold.pdf}
    \end{center}
    \caption{Performance of UTS-Mem with varying parallel loop
        threshol, run with 30 nodes, 6 cores per node, 4000 workers,
    \checkme{6M flush ticks}}
    \label{fig:uts_threshold}
\end{figure}
